---
title: "Check Your Setup"
subtitle: "MC 451 / MC 501 · Open this file first"
format: html
execute:
  eval: false
---

## What this file is

This is a **Quarto document**. It is a plain text file that holds two kinds of thing:
writing, like this paragraph, and **code chunks**, which are the grey boxes below. You
are reading it inside **VS Code**, the program we use for everything in this course.

You do not need to understand any of the code. You need to run it and see what happens.
That is the whole assignment.

Nothing in this file can break your computer, delete anything, or cost money.

## How to run a code chunk

Look at the grey box below. At its top left corner there is a small **Run Cell** link, or
a ▷ triangle. Click it.

If you do not see one, click anywhere inside the grey box and press **Ctrl+Enter**
(Windows) or **Cmd+Enter** (Mac). That sends the line your cursor is on to R.

The first time you do this, VS Code may take twenty or thirty seconds to wake R up, and a
new panel will open at the bottom of the screen. That panel is **the R console** — it is
where R answers you. Slowness the first time is normal and is not a sign that anything is
wrong.

## Step 1 — Is R alive?

This chunk asks R to say which version of itself it is running. It changes nothing. It is
a pulse check.

```{r}
R.version.string
```

**What you should see:** a line of text starting with `"R version 4."` and then some
numbers and a date.

**If you see that,** R is installed and VS Code can talk to it. Move on.

**If you see an error, or nothing happens at all,** stop here. Either R is not installed
or the R extension is not installed. Do not try to fix it by installing things at random —
write down exactly what the screen says and bring it to office hours or email it to me.
A screenshot is perfect.

## Step 2 — Install the helper that fetches our course package

Our course package does not live in R's usual store, so we need one small tool that knows
how to fetch it from somewhere else. That tool is called `remotes`.

You only ever run this once, on this computer, for the whole semester.

```{r}
install.packages("remotes")
```

**What you should see:** a lot of text scrolling past, some of it red, ending in something
like `package 'remotes' successfully unpacked`.

**Red text in R is not automatically an error.** R writes ordinary progress messages in
red. Only the last line matters. If R asks you a question in the console — for example
whether to use a "personal library" — answer **yes**.

## Step 3 — Install the course package

Now we fetch the course package itself, `v2v`. It holds our datasets and the shortcuts we
will use all term.

This one takes a few minutes the first time. Let it run. Go get water.

`upgrade = "never"` matters. Without it R offers to update dozens of other packages, and
saying yes to that on Windows fails halfway: Windows will not overwrite a file that another
program is holding open, so some packages roll back to the version you already had and the
install dies on the one it needed.

```{r}
remotes::install_github("AURA-Lab-SIUE/v2v-r", upgrade = "never")
```

**One thing that confuses everyone, so read it once now:** the thing we *download* is
called `v2v-r`, and the thing we *use* is called `v2v`. Two names, same course package.
That is normal for R and it is not a typo.

If R asks **"Which would you like to update?"** and offers a numbered list, type **1** and
press Enter. If it asks about installing from source, answer **no**.

## Step 4 — Switch the package on

Installing a package puts it on your computer. It does not switch it on. `library()`
switches it on, and you do this **every time you start a new R session** — not just once.

```{r}
library(v2v)
```

**What you should see:** nothing, or a short startup message. In R, silence usually means
success.

## Step 5 — The real check

This is the one that matters. `setup()` is a function we wrote for this course. It looks
over your whole installation and prints a checklist: green checks for what is working,
red marks for what is missing.

```{r}
setup()
```

**What you should see:** a list of package names with green check marks, then a
**Command-line tools** section.

**Two red marks are expected and fine at this stage** if you have not done the later
setup steps yet: Quarto CLI and Git. Everything under **Required R packages** should be
green. If a required package is red, `setup()` prints the exact command to fix it — copy
that command, paste it into the console, press Enter, then run `setup()` again.

## Step 6 — Prove it can actually get data

Last one. This pulls five rows of real Twitch chat out of the course dataset, so you can
see that the package is not just installed but working.

```{r}
twitch_chat(n = 5)
```

**What you should see:** a small table, five rows, with columns of usernames and message
text.

## You are done

If Step 6 gave you five rows of chat, your setup is finished and you are ready for class.
You do not need to do anything else.

**Please send me one thing:** a screenshot of the output of Step 5, `setup()`. That is how
I know you are ready, and it lets me spot a problem before it costs you a class period.

## If you got stuck

That is genuinely expected. Installing this stack breaks for someone in every section,
every semester, and it is never because that person is bad at this.

Do this, in order:

1. Close VS Code completely and reopen it. This fixes a surprising share of problems.
2. Run the chunk again. Some failures are a slow network, not a real fault.
3. Email me the **exact** text of the error, or a screenshot, plus which step number you
   were on. Do not paraphrase the error — the wording is the diagnosis.
4. Come to office hours: Tuesday and Thursday, 12:30 to 1:45 and 3:30 to 5:00, Dunham
   Hall 1017. Drop in, no appointment.

What I do not want is for you to spend two hours fighting this alone. Fifteen minutes,
then ask.
